---
title: "Biochemical Analysis 20 Nov 2017"
author: "Chandra Jack"
date: "November 20, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, warning=FALSE, message=FALSE}
library(ggplot2)
library(data.table)
library(scales)
library(plyr)
library(dplyr)
library(reshape2)
library(lme4)
library(effects)
library(multcomp)
library(lmerTest)
library(piecewiseSEM)
library(car)
library(gridExtra)
library(cowplot)
library(Rmisc)
library(heplots)
library(Hmisc)
library(RcmdrMisc)
library(corrplot)
```

This is the set of packages that I regularly use. Not all of them are necessary to run the code below. This is a compilation of the key parts of the analysis that has been done.

## Bootstrapping 

There are 11 Invasive genotypes representing 5 populations but 70 genotypes 13 populations for the Native range. I ran a Type II MANOVA for each assay instead of a repeated measures ANOVA due to the unbalanced design and because it is less sensitive to heterogeneity of variances. However, I still wanted to check to see if differences in group size could change significance. 

```{r, echo=FALSE}
BiocAssaysWide <- read.csv("~/Documents/Friesen lab/MedicagoHerbPopulation/HerbivoryCollabWSU/Data/ProcessedData/AllBiocAssaysMANOVADataStyle14Nov2017.csv") 
PopulationTable <- BiocAssaysWide %>% count(Range, Site)
PopulationTable$NoGenotypes <- PopulationTable$n/3
PopTab <- dcast(PopulationTable, Site ~ Range, value.var = "NoGenotypes")
ggplot(PopulationTable, aes(Site, n/3, fill = Range, label = n/3)) + geom_bar(stat = "identity") + ylab("Number of Genotypes") + theme(axis.text.x = element_text(angle = 45)) + geom_text()
```


### T test of bootstrap variance

I resampled from the native population, calcuated the variance and compared it to the variance from the full range set

Subset the native and invasive populations
```{r BootTSubset, warning=FALSE, message=FALSE}
BV_Nat <- subset(BiocAssaysWide, Range == "Native")
BV_Inv <- subset(BiocAssaysWide, Range == "Invasive")
```

```{r BootTRun, warning=FALSE, message= FALSE}
N = 1000 # Number of times to resample
TestMat <- matrix(rep(0, 9*N), ncol = 9) # Define the matrix to hold bootstrapped values and initiate with 0's
for (i in 1:N){
  pops <- sample(unique(BV_Nat$Genotype), 11) # Sample populations randomly keeping the rows and replicates together
  BV_Nat2 <- BV_Nat[BV_Nat$Genotype %in% pops, ] # Narrow dataset
  TestMat[i,] <- apply(BV_Nat2[,6:ncol(BV_Nat2)], 2, function(x) var(x, na.rm=TRUE)) # Calculate variance and fill matrix
}

VecVar <- apply(BV_Nat[,6:ncol(BV_Nat)], 2, function(x) var(x, na.rm=TRUE)) # Variance of all native genotypes
ColnNames <- colnames(BV_Nat)
ColnNames <- ColnNames[-(1:5)]

NewMat <- matrix(rep(0, 27), ncol = 3) # Place in its own matrix to call later
for (i in 1:9) {
  NewMat[i,1] <- ColnNames[i] # each variable tested
  NewMat[i,2] <- t.test(TestMat[,i], mu = VecVar[i])$p.value # p value comparing vector of bootstrap variances to observed variance
  NewMat[i,3] <- mean(abs(TestMat[,i] -mean(TestMat[,i])) > VecVar[i]) # a way to calculate a p value for the bootstrap
}
NewMat
```
According to this metric, the variances do not differ between the bootstrapped test and the observed variance.

### Levene's Test resample

I resampled the native range, combined it with the invasive range dataset and then ran the LeveneTest to compare variances
```{r BootLev}
N = 1000
TestMat <- matrix(rep(0, 9*N), ncol = 9)
for (i in 1:N){
  pops <- sample(unique(BV_Nat$Genotype), 11, replace = T) # Sample populations randomly 
  BV_Nat2 <- BV_Nat[BV_Nat$Genotype %in% pops, ] # Narrow dataset
  FullBV <- rbind(BV_Nat2, BV_Inv)
  test <- leveneTests(FullBV[,6:14], FullBV$Range)
  FStat <- test[,3] # Gets the F statistic from the Levene Test
  for (j in 1:9){
    TestMat[i, j] <- FStat[j]
  }
}

ColnNames <- colnames(BV_Nat)
ColnNames <- ColnNames[-(1:5)]
testObs <- leveneTests(BiocAssaysWide[,6:14], BiocAssaysWide$Range)
FStatObs <- testObs[,3] # Observed F statistic

FMat <- matrix(rep(0, 18), ncol = 2)
for (i in 1:9) {
  FMat[i,1] <- ColnNames[i]
  FMat[i,2] <- mean(abs(TestMat[,i] -mean(TestMat[,i])) > FStatObs[i]) # p value for bootstrapped difference between groups
}
FMat
```
Under this metric there is also not a significant difference between group variances


### Bootstrap univariate MANOVAs

 I resampled the native range and ran univariate MANOVAs. I'm going to run all of the code and then show the outputs together after.

```{r BootMANOVAGetData, include=FALSE}
PODWide <- read.csv("~/Documents/Friesen lab/MedicagoHerbPopulation/HerbivoryCollabWSU/Data/ProcessedData/PeroxidaselDataWideFormat25Oct2017.csv")
PPOWide <- read.csv("~/Documents/Friesen lab/MedicagoHerbPopulation/HerbivoryCollabWSU/Data/ProcessedData/PolyphenolDataWideFormat25Oct2017.csv")
ProWide <- read.csv("/Users/chanj/Documents/Friesen\ lab/MedicagoHerbPopulation/HerbivoryCollabWSU/Data/ProcessedData/ProteinWideFormat25Oct2017.csv")
```

This section of code is also used to run the repeated measures MANOVA on the full dataset
```{r RepeatedMeasuresMANOVA}
block <- ordered(c("Hr0", "Hr4", "Hr24"), levels = c("Hr0", "Hr4", "Hr24"))
contrasts(block) <- matrix(c(-1,1,0,0,-1,1), ncol = 2)
idata <- data.frame(block)
```

#### Bootstrap Peroxidase Assay
```{r bootPod}
PODWide_Nat <- subset(PODWide, Range == "Native")
PODWide_Inv <- subset(PODWide, Range == "Invasive")

N = 1000
PodMat <- matrix(rep(0, 6*N), ncol = 6)

for (i in 1:N) {
  outtests <- car:::print.Anova.mlm
  body(outtests)[[16]] <- quote(invisible(tests))
  body(outtests)[[15]] <- NULL
  pops <- sample(unique(PODWide_Nat$Genotype), 11, replace = T) 
  PODWide_Nat2 <- PODWide_Nat[PODWide_Nat$Genotype %in% pops, ] # Narrow dataset
  FullPODWide <- rbind(PODWide_Nat2, PODWide_Inv)
  PodMod1 <- lm(cbind(PODHr0, PODHr4, PODHr24) ~ Range, data = FullPODWide)
  tab <- lapply(c("Pillai", "Wilks", "Hotelling-Lawley", "Roy"),  function(i)  outtests(Anova(PodMod1, idata = idata, idesign = ~ block, test.statistic=i)))
  tab <- do.call(rbind, tab)
  Fval <- tab[,3]
  Pval <- tab[,6]
  for (j in 1:3){
    X = j + 1
    Y = j + 3
    PodMat[i,j] <- Fval[X]
    PodMat[i,Y] <- Pval[X]
  }
}
```

#### Bootstrap Polyphenol Oxidase Assay
```{r bootPPO}
PPOWide_Nat <- subset(PPOWide, Range == "Native")
PPOWide_Inv <- subset(PPOWide, Range == "Invasive")
N = 1000
PpoMat <- matrix(rep(0, 6*N), ncol = 6)

for (i in 1:N) {
  outtests <- car:::print.Anova.mlm
  body(outtests)[[16]] <- quote(invisible(tests))
  body(outtests)[[15]] <- NULL
  pops <- sample(unique(PPOWide_Nat$Genotype), 11, replace = T) 
  PPOWide_Nat2 <- PPOWide_Nat[PPOWide_Nat$Genotype %in% pops, ] # Narrow dataset
  FullPPOWide <- rbind(PPOWide_Nat2, PPOWide_Inv)
  PpoMod1 <- lm(cbind(PPOHr0, PPOHr4, PPOHr24) ~ Range, data = FullPPOWide)
  tab <- lapply(c("Pillai", "Wilks", "Hotelling-Lawley", "Roy"),  function(i)  outtests(Anova(PpoMod1, idata = idata, idesign = ~ block, test.statistic=i)))
  tab <- do.call(rbind, tab)
  Fval <- tab[,3]
  Pval <- tab[,6]
  for (j in 1:3){
    X = j + 1
    Y = j + 3
    PpoMat[i,j] <- Fval[X]
    PpoMat[i,Y] <- Pval[X]
  }
}
```

#### Bootstrap Protein Quantification Assay
```{r}
ProWide_Nat <- subset(ProWide, Range == "Native")
ProWide_Inv <- subset(ProWide, Range == "Invasive")
N = 1000
ProMat <- matrix(rep(0, 6*N), ncol = 6)

for (i in 1:N) {
  outtests <- car:::print.Anova.mlm
  body(outtests)[[16]] <- quote(invisible(tests))
  body(outtests)[[15]] <- NULL
  pops <- sample(unique(PODWide_Nat$Genotype), 11, replace = T) 
  ProWide_Nat2 <- ProWide_Nat[ProWide_Nat$Genotype %in% pops, ] # Narrow dataset
  FullProWide <- rbind(ProWide_Nat2, ProWide_Inv)
  ProMod1 <- lm(cbind(ProteinHr0, ProteinHr4, ProteinHr24) ~ Range, data = FullProWide)
  tab <- lapply(c("Pillai", "Wilks", "Hotelling-Lawley", "Roy"),  function(i)  outtests(Anova(ProMod1, idata = idata, idesign = ~ block, test.statistic=i)))
  tab <- do.call(rbind, tab)
  Fval <- tab[,3]
  Pval <- tab[,6]
  for (j in 1:3){
    X = j + 1
    Y = j + 3
    ProMat[i,j] <- Fval[X]
    ProMat[i,Y] <- Pval[X]
  }
}
```

This gives the observed output. It is arranged by F statistic of the observed model, p value of the observed model, and p value for the F statistic of the bootstrap
```{r}
#pod
outtests <- car:::print.Anova.mlm
body(outtests)[[16]] <- quote(invisible(tests))
body(outtests)[[15]] <- NULL
PodMod1 <- lm(cbind(PODHr0, PODHr4, PODHr24) ~ Range, data = PODWide)
tab <- lapply(c("Pillai", "Wilks", "Hotelling-Lawley", "Roy"),  function(i)  outtests(Anova(PodMod1, idata = idata, idesign = ~ block, test.statistic=i)))
tab <- do.call(rbind, tab)

Fval <- tab[,3]
Pval <- tab[,6]
PODtrue <- data.frame(Fval[2:4], Pval[2:4])
rownames(PODtrue) <- c("Range", "block", "Range:block")
names(PODtrue)[names(PODtrue) == 'Fval.2.4.'] <- 'test.statistic'
names(PODtrue)[names(PODtrue) == 'Pval.2.4.'] <- 'p.value'

for (i in 1:3) {
  PODtrue$BootF[i] <-mean(abs(PodMat[,i] - mean(PodMat[,i])) > PODtrue$test.statistic[i])
}

# PPO
outtests <- car:::print.Anova.mlm
body(outtests)[[16]] <- quote(invisible(tests))
body(outtests)[[15]] <- NULL
PpoMod1 <- lm(cbind(PPOHr0, PPOHr4, PPOHr24) ~ Range, data = PPOWide)
tab <- lapply(c("Pillai", "Wilks", "Hotelling-Lawley", "Roy"),  function(i)  outtests(Anova(PpoMod1, idata = idata, idesign = ~ block, test.statistic=i)))
tab <- do.call(rbind, tab)

Fval <- tab[,3]
Pval <- tab[,6]
PPOtrue <- data.frame(Fval[2:4], Pval[2:4])
rownames(PPOtrue) <- c("Range", "block", "Range:block")
names(PPOtrue)[names(PPOtrue) == 'Fval.2.4.'] <- 'test.statistic'
names(PPOtrue)[names(PPOtrue) == 'Pval.2.4.'] <- 'p.value'

for (i in 1:3) {
  PPOtrue$BootF[i] <-mean(abs(PpoMat[,i] - mean(PpoMat[,i])) > PPOtrue$test.statistic[i])
}      

# PROTEIN
outtests <- car:::print.Anova.mlm
body(outtests)[[16]] <- quote(invisible(tests))
body(outtests)[[15]] <- NULL
ProMod1 <- lm(cbind(ProteinHr0, ProteinHr4, ProteinHr24) ~ Range, data = ProWide)
tab <- lapply(c("Pillai", "Wilks", "Hotelling-Lawley", "Roy"),  function(i)  outtests(Anova(ProMod1, idata = idata, idesign = ~ block, test.statistic=i)))
tab <- do.call(rbind, tab)

Fval <- tab[,3]
Pval <- tab[,6]
PROtrue <- data.frame(Fval[2:4], Pval[2:4])
rownames(PROtrue) <- c("Range", "block", "Range:block")
names(PROtrue)[names(PROtrue) == 'Fval.2.4.'] <- 'test.statistic'
names(PROtrue)[names(PROtrue) == 'Pval.2.4.'] <- 'p.value'

for (i in 1:3) {
  PROtrue$BootF[i] <-mean(abs(ProMat[,i] - mean(ProMat[,i])) > PROtrue$test.statistic[i])
}      

PODtrue
PPOtrue
PROtrue
```
p values for the bootstrapped values seem elevated compared to the observed value. 


## Repeated measures MANOVAs for each assay.

Diagnostic steps for these assays can be found in the most Induced vs constitutive pdf. 

### Peroxidase

1. Does the invasive range differ in POD activity significantly from the native range?

2. Do the two ranges respond similarly over time?
```{r}
PodMod1 <- lm(cbind(PODHr0, PODHr4, PODHr24) ~ Range, data = PODWide)
Anova(PodMod1, idata = idata, idesign = ~ block)
```
Range is not significant but there are differences between the two over time.


### Polyphenol Oxidase

1. Does the invasive range differ in PPO activity significantly from the native range?

2. Do the two ranges respond similarly over time?
```{r}
PpoMod1 <- lm(cbind(PPOHr0, PPOHr4, PPOHr24) ~ Range, data = PPOWide)
Anova(PpoMod1, idata = idata, idesign = ~ block)
```
PPO activity varies between the ranges but both respond similarly over time


### Protein Quantification

1. Does the invasive range differ in protein production significantly from the native range?

2. Do the two ranges respond similarly over time?
```{r}
ProMod1 <- lm(cbind(ProteinHr0, ProteinHr4, ProteinHr24) ~ Range, data = ProWide)
Anova(ProMod1, idata = idata, idesign = ~ block)
```
Protein production varies between the ranges but both respond similarly over time

### Plot of assay results

```{r, echo=FALSE, warning= FALSE, message= FALSE}
ProteinUntra <- read.csv("/Users/chanj/Documents/Friesen lab/MedicagoHerbPopulation/HerbivoryCollabWSU/Data/ProcessedData/ProteinData22Feb2017.csv")

PODUntra <- read.csv("/Users/chanj/Documents/Friesen lab/MedicagoHerbPopulation/HerbivoryCollabWSU/Data/ProcessedData/WSU_PODfiles11Sept2017.csv")

PPOUntra <- read.csv("/Users/chanj/Documents/Friesen\ lab/MedicagoHerbPopulation/HerbivoryCollabWSU/Data/ProcessedData/WSU_PPOfiles11Sept2017.csv")

# Remove all extraneous columns
ProteinUnneeded <- names(ProteinUntra) %in% c("Well", "Weight", "Hour", "PCheck", "Plate", "fileName", "Absorbance")
ProteinUntra <- ProteinUntra[!ProteinUnneeded]

PODKeep <- names(PODUntra) %in% c("Sample", "Replicate", "Time", "AbsFreshWeight", "Range", "Genotype", "Site")
PODUntra <- PODUntra[PODKeep]

PPOKeep <- names(PPOUntra) %in% c("Sample", "Replicate", "Time", "AbsFreshWeight", "Range", "Genotype", "Site")
PPOUntra <- PPOUntra[PPOKeep]

# Rename response variable so that can bind the rows
names(ProteinUntra)[names(ProteinUntra) == "Protein"] <- "Response"
names(PODUntra)[names(PODUntra) == "AbsFreshWeight"] <- "Response"
names(PPOUntra)[names(PPOUntra) == "AbsFreshWeight"] <- "Response"

# Add column that identifies each assay
ProteinUntra$Assay <- rep("Protein", nrow(ProteinUntra))
PODUntra$Assay <- rep("POD", nrow(PODUntra))
PPOUntra$Assay <- rep("PPO", nrow(PPOUntra))

# Combine datasets
AllBiocUntrans <- rbind(ProteinUntra, PODUntra)
AllBiocUntrans <- rbind(AllBiocUntrans, PPOUntra)

ABU <- na.omit(AllBiocUntrans)
ggplot(ABU, aes(x = as.factor(Time), y = Response, shape = Range, colour = Range)) + stat_summary(fun.data = "mean_se") + facet_wrap(~Assay, scales = "free") + stat_summary(fun.y = mean, geom = "line", aes(group = Range))
```

### Correlations between all the assays and times

A pearson correlation plot where the intensity of the color and size of the circle indicate correlations. Blue are positive and Red are negative. Only significant correlations are shown (Holm's adjusted p values < 0.05)
```{r, echo = FALSE, warning = FALSE}
CorData <- BiocAssaysWide[, c(6:14)]
CDmatrix3 <- rcorr.adjust(as.matrix(CorData))
corrplot(CDmatrix3$R$r, type = "upper", p.mat = apply(CDmatrix3$P, 2, as.numeric)
, sig.level = 0.05, insig = "blank")
```


