---
title: 'Induced vs Constitutive Project: Univariate repeated analysis of biochemical
  assays'
author: "Chandra Jack"
date: "October 27, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Results of the univariate analyses for each biochemical assay

Prior to running the multivariate analysis, I looked at each assay separately as a repeated measures test.

```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(data.table)
library(scales)
library(plyr)
library(dplyr)
library(reshape2)
library(lme4)
library(effects)
library(multcomp)
library(lmerTest)
library(piecewiseSEM)
library(car)
library(gridExtra)
library(cowplot)
library(Rmisc)
```

## Protein Quantification Analysis

Boxplots to examine for homogeneity of variances and density plots for normal distributions

```{r pressure, echo=FALSE, warning=FALSE}
ProteinFile <- read.csv("~/Documents/Friesen lab/MedicagoHerbPopulation/HerbivoryCollabWSU/Data/ProcessedData/ProteinData22Feb2017.csv")

# Removed unnecessary columns
ProRemove <- names(ProteinFile) %in% c("PCheck", "Plate", "fileName", "Absorbance")
ProteinFile <- ProteinFile[!ProRemove]

# Removed USDA genotypes
MPmaster <- read.csv("~/Documents/Friesen lab/MedicagoHerbPopulation/HerbivoryCollabWSU/Copy of WSUV_MpolMlup_BulkingDesign_3.2016_masterfile -1.csv")

MP_USDA <- MPmaster$Unique.ID[MPmaster$Pod.Produced != "USDA"]

ProteinFile <- ProteinFile[ProteinFile$Sample %in% MP_USDA,]

ProteinFile$Hour <- factor(ProteinFile$Hour, levels = c("0hr", "4hr", "24hr"))

P1 <-ggplot(ProteinFile, aes(Hour, Protein, fill = Range)) + geom_boxplot()
P2 <- ggplot(ProteinFile, aes(Hour, Protein, fill = Range)) + geom_boxplot() +ylim(0, 1*10^6)
ProBoxplot <-ggdraw() + draw_plot(P1 + theme(legend.position = "none"), 0, 0, 1, 1) + draw_plot(P2, 0.5, 0.6, 0.5, 0.5)
ProBoxplot

ProteinFile$ProSqrt <- sqrt(ProteinFile$Protein)
ProteinFile$ProCube <- ProteinFile$Protein ^ 0.33
ProteinFile$ProLog <- log(ProteinFile$Protein)

opar <- par()      # make a copy of current settings
par(mfrow=c(2,2))
densityPlot(ProteinFile$Protein)
densityPlot(ProteinFile$ProSqrt)
densityPlot(ProteinFile$ProCube)
densityPlot(ProteinFile$ProLog)
par(opar)
```

Results of the univariate repeated measures

```{r protein uni, echo=FALSE, warning=FALSE}
ProteinWide <- dcast(ProteinFile, Sample + Site + Replicate + Genotype + Range ~ Hour, value.var = "ProLog")
ProteinWide <- na.exclude(ProteinWide)
colnames(ProteinWide)[colnames(ProteinWide) =="0hr"] <- "ProteinHr0"
colnames(ProteinWide)[colnames(ProteinWide) =="4hr"] <- "ProteinHr4"
colnames(ProteinWide)[colnames(ProteinWide) =="24hr"] <- "ProteinHr24"
```

```{r}
ProMod1 <- lm(cbind(ProteinHr0, ProteinHr4, ProteinHr24) ~ Range, data = ProteinWide)
```

Residuals:
```{r, echo= FALSE}
plot(density(resid(ProMod1)))
```

```{r, echo=FALSE}
blockD <- ordered(c("Hr0", "Hr4", "Hr24"), levels = c("Hr0", "Hr4", "Hr24"))
contrasts(blockD) <- matrix(c(-1,1,0,0,-1,1), ncol = 2)
idata <-data.frame(blockD)

proMod1a <-Anova(ProMod1, idata = idata, idesign = ~ blockD)
proMod1a
```
## Peroxidase Activity Assay

```{r, echo=F, warning=F}
PODfile <- read.csv("~/Documents/Friesen lab/MedicagoHerbPopulation/HerbivoryCollabWSU/Data/ProcessedData/WSU_PODfiles11Sept2017.csv")

# Removed unnecessary columns
PODRemove <- names(PODfile) %in% c("PCheck", "Plate", "fileName", "values", "MatchValue", "CtrlVal","ABStrt_ABSctr")
PODfile <- PODfile[!PODRemove]

PODfile <- PODfile[PODfile$Sample %in% MP_USDA,]
PODfile$Hour <- factor(PODfile$Hour, levels = c("0hr", "4hr", "24hr"))
Pd1<- ggplot(PODfile, aes(Hour, AbsFreshWeight, fill = Range)) + geom_boxplot() 
Pd2 <- ggplot(PODfile, aes(Hour, AbsFreshWeight, fill = Range)) + geom_boxplot() +ylim(0, 1*10^2)
ProBoxplot <-ggdraw() + draw_plot(Pd1 + theme(legend.position = "none"), 0, 0, 1, 1) + draw_plot(Pd2, 0.1, 0.5, 0.5, 0.5)
ProBoxplot

PODfile$AbsFreshWeightN <- PODfile$AbsFreshWeight + 4
PODfile$PodSqrt <- sqrt(PODfile$AbsFreshWeightN)
PODfile$PodCube <- PODfile$AbsFreshWeightN ^ 0.33
PODfile$PodLog <- log(PODfile$AbsFreshWeightN)
par(mfrow=c(2,2))
densityPlot(PODfile$AbsFreshWeightN)
densityPlot(PODfile$PodSqrt)
densityPlot(PODfile$PodCube)
densityPlot(PODfile$PodLog)
par(opar)
```

These looked a little odd so I broke them down by range. Graph on the right is at same scale, left is to give better view of distribution

```{r, echo=FALSE, warning= F}
F1 <-ggplot(PODfile, aes(PodLog, fill = Range)) + geom_density()  + facet_grid(Range~ Hour, scales = "free")
F2 <-ggplot(PODfile, aes(PodLog, fill = Range)) + geom_density()  + facet_grid(Range~ Hour)
PODdist <-plot_grid(F1, F2)
PODdist
```

Results of the univariate repeated measures
```{r pod uni, echo=FALSE, warning=FALSE}
PODWide <- dcast(PODfile, Sample + Site + Replicate + Genotype + Range ~ Hour, value.var = "PodLog")
PODWide <- na.exclude(PODWide)
colnames(PODWide)[colnames(PODWide) =="0hr"] <- "PODHr0"
colnames(PODWide)[colnames(PODWide) =="4hr"] <- "PODHr4"
colnames(PODWide)[colnames(PODWide) =="24hr"] <- "PODHr24"
```

```{r}
PodMod1 <- lm(cbind(PODHr0, PODHr4, PODHr24) ~ Range, data = PODWide)
```

Residuals:
```{r, echo= FALSE}
plot(density(resid(PodMod1)))
```

```{r, echo= F}
Anova(PodMod1, idata = idata, idesign = ~ blockD)
```

## Polyphenol Oxidase Activity

```{r, echo=FALSE, warning=FALSE}
PPOFile <- read.csv("~/Documents/Friesen lab/MedicagoHerbPopulation/HerbivoryCollabWSU/Data/ProcessedData/WSU_PPOfiles11Sept2017.csv")
PPOFile <- subset(PPOFile, Sample !="")

# Removed unnecessary columns
ProRemove <- names(PPOFile) %in% c("PCheck", "Plate", "fileName", "values", "MatchValue", "CtrlVal","ABStrt_ABSctr")
PPOFile <- PPOFile[!ProRemove]
PPOFile <- PPOFile[PPOFile$Sample %in% MP_USDA,]
PPOFile$Hour <- factor(PPOFile$Hour, levels = c("0hr", "4hr", "24hr"))

ggplot(PPOFile, aes(Hour, AbsFreshWeight, fill = Range)) + geom_boxplot() 
PPOFile$AbsFreshWeightN <- PPOFile$AbsFreshWeight + 1
PPOFile$PpoSqrt <- sqrt(PPOFile$AbsFreshWeightN)
PPOFile$PpoCube <- PPOFile$AbsFreshWeightN ^ 0.33
PPOFile$PpoLog <- log(PPOFile$AbsFreshWeightN)
par(mfrow=c(2,2))
densityPlot(PPOFile$AbsFreshWeightN)
densityPlot(PPOFile$PpoSqrt)
densityPlot(PPOFile$PpoCube)
densityPlot(PPOFile$PpoLog)
par(opar)

ggplot(PPOFile, aes(PpoLog, fill = Range)) + geom_density()  + facet_grid(Range~ Hour)
```

Results of the univariate repeated measures

```{r, echo=FALSE, warning=FALSE}
PPOWide <- dcast(PPOFile, Sample + Site + Replicate + Genotype + Range ~ Hour, value.var = "PpoLog")
PPOWide <- na.exclude(PPOWide)
colnames(PPOWide)[colnames(PPOWide) =="0hr"] <- "PPOHr0"
colnames(PPOWide)[colnames(PPOWide) =="4hr"] <- "PPOHr4"
colnames(PPOWide)[colnames(PPOWide) =="24hr"] <- "PPOHr24"
```

```{r}
PPOMod1 <- lm(cbind(PPOHr0, PPOHr4, PPOHr24) ~ Range, data = PPOWide)
```

Residuals:
```{r, echo= FALSE}
plot(density(resid(PPOMod1)))
```

```{r echo=FALSE}
Anova(PPOMod1, idata = idata, idesign = ~ blockD)
```

