---
title: "Multivariate analysis of biochemical assayss"
author: "Chandra Jack"
date: "November 6, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Multivariate analysis of biochemical assayss



```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(data.table)
library(scales)
library(plyr)
library(dplyr)
library(reshape2)
library(lme4)
library(effects)
library(multcomp)
library(lmerTest)
library(piecewiseSEM)
library(car)
library(gridExtra)
library(cowplot)
library(Rmisc)
library(Hmisc)
library(heplots)
library(corrplot)
library(PerformanceAnalytics)
```


```{r echo=FALSE, warning=FALSE}
setwd("~/Documents/Friesen\ lab/MedicagoHerbPopulation/HerbivoryCollabWSU/Data/ProcessedData/")

ProteinWide  <- read.csv("/Users/chanj/Documents/Friesen\ lab/MedicagoHerbPopulation/HerbivoryCollabWSU/Data/ProcessedData/ProteinWideFormat25Oct2017.csv")

PODWide <- read.csv("/Users/chanj/Documents/Friesen\ lab/MedicagoHerbPopulation/HerbivoryCollabWSU/Data/ProcessedData/PeroxidaselDataWideFormat25Oct2017.csv")

PPOWide <- read.csv("/Users/chanj/Documents/Friesen\ lab/MedicagoHerbPopulation/HerbivoryCollabWSU/Data/ProcessedData/PolyphenolDataWideFormat25Oct2017.csv")

BiocAssaysWide <- merge(ProteinWide, PODWide)
BiocAssaysWide <- merge(BiocAssaysWide, PPOWide)

# Time for the big jump! 

Assay <- factor(rep(c("Protein", "POD", "PPO"), each = 3))
MeasTime <- factor(rep(c("Hr0", "Hr4", "Hr24"), 3), levels = c("Hr0", "Hr4", "Hr24"))

idata <- data.frame(Assay, MeasTime)

```

``` {r pressure, warning=FALSE}
BCmod1 <- lm(cbind(ProteinHr0, ProteinHr4, ProteinHr24, PODHr0, PODHr4, PODHr24,PPOHr0, PPOHr4, PPOHr24) ~ Range, data = BiocAssaysWide)

BCmod2 <- lm(cbind(ProteinHr0, ProteinHr4, ProteinHr24, PODHr0, PODHr4, PODHr24,PPOHr0, PPOHr4, PPOHr24) ~ Site, data = BiocAssaysWide)

anova.BCmod1 <- Anova(BCmod1, idata= idata, idesign = ~ Assay * MeasTime, iterm = "Assay:MeasTime", type = 3)
anova.BCmod2 <- Anova(BCmod2, idata= idata, idesign = ~ Assay * MeasTime,iterm = "Assay:MeasTime", type = 3)

```

There is a significant effect for Population, but not for Range

``` {r, echo = FALSE, warning = FALSE}
anova.BCmod1
heplot(BCmod1)

anova.BCmod2
heplot(BCmod2)
```

## Correlation Data

Plot 1: Correlation plot
Plot 2: Keeping only dots that are significant at 0.05
Plot 3: Numbers instead of dot size
Plot 4: All the info

``` {r, echo = FALSE,  warning = FALSE}
CorData <- BiocAssaysWide[, c(6:14)]
CDmatrix <- cor(CorData)
CDmatrix2 <- rcorr(as.matrix(CorData))

corrplot(CDmatrix, type = "upper")
corrplot(CDmatrix2$r, type = "upper", p.mat = CDmatrix2$P, sig.level = 0.05, insig = "blank")
corrplot(CDmatrix2$r, type = "upper", p.mat = CDmatrix2$P, sig.level = 0.05, insig = "blank", method = "number")
chart.Correlation(CorData, histogram=TRUE, pch=19)
