---
title: "Choice Data ANOVA tables"
author: "Chandra Jack"
date: "December 11, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(data.table)
library(scales)
library(plyr)
library(dplyr)
library(reshape2)
library(lme4)
library(effects)
library(multcomp)
library(lmerTest)
library(piecewiseSEM)
library(car)
library(gridExtra)
library(cowplot)
library(Rmisc)
library(heplots)
library(Hmisc)
library(RcmdrMisc)
library(corrplot)

setwd("~/Documents/Friesen lab/MedicagoHerbPopulation/HerbivoryCollabWSU/Data/ProcessedData/ChoiceExpts/")

PrefData <- read.csv("Corrected_FINAL_PreferenceData.csv")
MPmaster <- read.csv("~/Documents/Friesen lab/MedicagoHerbPopulation/HerbivoryCollabWSU/Data/ProcessedData/WSUV_BulkingData_Final_ZCL.csv")

# Preparing the dataset
PrefData$Genotype_Plant1 <- MPmaster[match(PrefData$PlantID_1, MPmaster$Unique.ID), "MSU.Genotype"]
PrefData$Population_Plant1 <- MPmaster[match(PrefData$PlantID_1, MPmaster$Unique.ID), "Site.ID"]

PrefData$Genotype_Plant2 <- MPmaster[match(PrefData$PlantID_2, MPmaster$Unique.ID), "MSU.Genotype"]
PrefData$Population_Plant2 <- MPmaster[match(PrefData$PlantID_2, MPmaster$Unique.ID), "Site.ID"]

PrefData$TissuePref <- PrefData$AmtPlant1Consumed - PrefData$AmtPlant2Consumed

```

## Preference Data: Constitutive defenses

The script used for this is the PrefPlastic17Nov2017.R file found in the src folder https://drive.google.com/open?id=14cdMdq77dwGCwBEXXVfOB7uV_1m1BdBV

The dataset is found in the constitutive data folder
https://drive.google.com/open?id=1RRla12LJq7zppEN5uRPZvezYLPzLg251

I've uploaded the excel spreadsheet so that you can see the formulas that I used outside of R. You can convert that to a csv file if you want to follow from the beginning, otherwise use the uploaded modified csv file that starts at line 36.

Running a glm binomial model to check for preference for invasive or native range where 1 is invasive and 0 is native

```{r }
mod1 <- glm(as.factor(Inv_1_Nat_0) ~ Herbivore*CatStage , family = binomial(link = "logit"), data = PrefData)
anova(mod1, test = "Chisq")
```

I realized that I never plotted this data.

```{r constit, echo=FALSE}
ggplot(PrefData, aes(CatStage, Inv_1_Nat_0, colour = Herbivore)) + stat_summary(fun.data = "mean_se") + geom_hline(yintercept = 0.5, colour = "black")
```

From this it looks as if there should be some type of significant difference, at least in the soybean looper data, but from the above ANOVA, it doesn't show up. I am still concerned that there is a lack of replication but I subsetted the data and ran a non-parametric one sample t test (lines 43-48)

```{r}
PrefLoop <- subset(PrefData, Herbivore == "Soybean Looper")
PrefV <- subset(PrefData, Herbivore != "Soybean Looper")

# Looper data
wilcox.test(PrefLoop$Inv_1_Nat_0, mu = 0.5)
# Velvet bean data
wilcox.test(PrefV$Inv_1_Nat_0, mu = 0.5)
```


Here is the ANOVA table based on the amount actually consumed and the non-p t test results.

```{r echo= FALSE}

ggplot(PrefData, aes(CatStage, TissuePref, colour = Herbivore)) + stat_summary(fun.data = "mean_se") + geom_hline(yintercept = 0, colour = "black")
```

```{r}
mod3 <- lm(TissuePref ~ Herbivore*CatStage, data = PrefData)
anova(mod3)
M3.res <- rstandard(mod3)
qqnorm(M3.res)
qqline(M3.res)

# Looper data
wilcox.test(PrefLoop$TissuePref, mu = 0)
# Velvet bean data
wilcox.test(PrefV$TissuePref, mu = 0)
```


## Preference Data: Inducible/Plastic 

This is included in the same script. The dataset is found in the inducible folder
https://drive.google.com/open?id=1cY6FTjSg0KPTTiBgWsUsB8RI6H7Ezxm3

The excel spreadsheet is not in the folder just the csv file(s) I used. Use the Modified file to start at line 77.

```{r echo= FALSE}
PlasticData <- read.csv("Corrected_FINAL_PlasticityData.csv")

# Preparing the dataset
PlasticData$Genotype_Plant1 <- MPmaster[match(PlasticData$PlantID_1, MPmaster$Unique.ID), "MSU.Genotype"]
PlasticData$Population_Plant1 <- MPmaster[match(PlasticData$PlantID_1, MPmaster$Unique.ID), "Site.ID"]

PlasticData$Genotype_Plant2 <- MPmaster[match(PlasticData$PlantID_2, MPmaster$Unique.ID), "MSU.Genotype"]
PlasticData$Population_Plant2 <- MPmaster[match(PlasticData$PlantID_2, MPmaster$Unique.ID), "Site.ID"]

write.table(PlasticData, file = "ModifiedPlasticity7Dec2017cj.csv", sep = ",", row = F)
PlasticData$TissuePref <- PlasticData$AmtPlant1Consumed - PlasticData$AmtPlant2Consumed
```

Running a glm binomial model to check for preference where 1 is constitutive and 0 is invasive. I also included Range as a variable.

```{r}
mod1 <- glm(as.factor(Con_1_Ind_0) ~ Herbivore*CatStage + Range, family = binomial(link = "logit"), data = PlasticData)
anova(mod1, test = "Chisq")
```

```{r echo= FALSE}
ggplot(PlasticData, aes(CatStage, Con_1_Ind_0, colour = Herbivore)) + stat_summary(fun.data = "mean_se") + geom_hline(yintercept = 0.5, colour = "black") + facet_wrap(~ Range)
PlasticLoop <- subset(PlasticData, Herbivore == "Soybean Looper")
PlasticV <- subset(PlasticData, Herbivore != "Soybean Looper")
```

For completion, also included 2 group Mann-Whitney U
```{r}
# Looper
wilcox.test(PlasticLoop$Con_1_Ind_0 ~ PlasticLoop$Range)
#Velvet bean
wilcox.test(PlasticV$Con_1_Ind_0 ~ PlasticV$Range)
```


Here is the data based on the amount of tissue consumed 
```{r echo= FALSE}

ggplot(PlasticData, aes(CatStage, TissuePref, colour = Herbivore)) + stat_summary(fun.data = "mean_se") + geom_hline(yintercept = 0, colour = "black") + facet_wrap(~ Range)
```

```{r}
mod4 <- lm(TissuePref ~ Herbivore*CatStage + Range, data = PlasticData)
anova(mod4)
```

```{r}
# Looper
wilcox.test(PlasticLoop$TissuePref ~ PlasticLoop$Range)
#Velvet bean
wilcox.test(PlasticV$TissuePref ~ PlasticV$Range)
```
